---
title: "Week5 Lab"
author: "Elmera Azadpour"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(tsibble)
library(feasts)
library(forecast)
library(fable)

# Spatial data packages
library(sf)
library(tmap)
```

## Read in data
```{r}
energy <- read_csv(here("data", "energy.csv"))
```

## convert month from character to date using tsibble
```{r}
energy_ts <- energy %>% 
  mutate(date = yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)
```

## lets look at our time series data
```{r}
ggplot(energy_ts, aes(x = date, y = res_total)) +
  geom_line() +
  labs(y = "residential energy consumption \n (Trillion BTU)")
```

Trends: 
- increasing overall but slightly decreasing around 2005
- seasonality in peaks
- no cyclicality or outliers


## Seasonplot:
```{r}
energy_ts %>% 
  gg_season(y= res_total) +
  theme_minimal() +
  labs( x = "month",
        y= "residential energy consumption (trillion BTU)")
```
Trends:
- highest in winter months (dec, jan, feb)
- peaks back up around jul and aug
- increasing over time in general

## Subseries plot
```{r}
energy_ts %>% gg_subseries(res_total)
```
Trends:
- clear seasonailty with increasing trend overtime
- peaks in dec, jan, feb confirmed

## Decomposition
```{r}
# Find STL decomposition
dcmp <- energy_ts %>%
  model(STL(res_total ~ season()))

# View the components
components(dcmp)

# Visualize the decomposed components
components(dcmp) %>% autoplot() +
  theme_minimal()
```
## Autocorrelation function (ACF) 
```{r}
energy_ts %>% 
  ACF(res_total) %>% 
  autoplot()
```
Takeaway:
- observations separated by 12 months are the most highly correlated, reflecting strong seasonality we see in all of our other exploratory visualizations.

## D. Forecasting by Holt-Winters exponential smoothing
- model specifications:
 - "N" = non-seasonal 
 - "A" = additive
 - "M" = multiplicative
 
 
We'll use multiplicative due to changes in variance over time
```{r}
# Create the model:
energy_fit <- energy_ts %>%
  model(ets = ETS(res_total ~ season("M")))

# Forecast using the model 10 years into the future:
energy_forecast <- energy_fit %>% 
  forecast(h = "10 years")

# Plot just the forecasted values (with 80 & 95% CIs):
energy_forecast %>% 
  autoplot()
```
```{r}
# Or plot it added to the original data:
energy_forecast %>% 
  autoplot(energy_ts)
```
## Assessing residuals
```{r}
# Append the predicted values (and residuals) to original energy data
energy_predicted <- broom::augment(energy_fit)

# Use View(energy_predicted) to see the resulting data frame
```

Now, plot the actual energy values (res_total), and the predicted values (stored as .fitted) atop them:
```{r}
ggplot(data = energy_predicted) +
  geom_line(aes(x = date, y = res_total)) +
  geom_line(aes(x = date, y = .fitted), color = "red")
```
## Explore the residuals
```{r}
ggplot(data = energy_predicted, aes(x = .resid)) +
  geom_histogram()
```

## Extra forecasting bits
```{r}
# Fit 3 different forecasting models (ETS, ARIMA, SNAIVE):
energy_fit_multi <- energy_ts %>%
  model(
    ets = ETS(res_total ~ season("M")),
    arima = ARIMA(res_total),
    snaive = SNAIVE(res_total)
  )

# Forecast 3 years into the future (from data end date)
multi_forecast <- energy_fit_multi %>% 
  forecast(h = "3 years")

# Plot the 3 forecasts
multi_forecast %>% 
  autoplot(energy_ts)
```

```{r}
# Or just view the forecasts (note the similarity across models):
multi_forecast %>% 
  autoplot()
```

### Part 2: Spatial data wrangling, visualization and a variogram

A. California county outlines (polygons)
```{r}
ca_counties <- read_sf(here("data","ca_counties","CA_Counties_TIGER2016.shp"))
```

bit of wrangling 
```{r}
ca_subset <- ca_counties %>% 
  select(NAME, ALAND) %>% 
  rename(county_name = NAME, land_area = ALAND)
```

## Check and set the CRS

Use st_crs() to check the existing CRS for spatial data. We see that this CRS is WGS84 (epsg: 3857).
```{r}
ca_subset %>% st_crs()
```
```{r}
ggplot(data = ca_subset) +
  geom_sf(aes(fill = land_area), color = "white", size = 0.1) +
  theme_void() +
  scale_fill_gradientn(colors = c("cyan","blue","purple"))
```
B. Invasive red sesbania records (spatial points)

- The data exist data/red_sesbania, and the shapefile is stored as ds80.shp. Let’s read in the data:
```{r}
sesbania <- read_sf(here("data","red_sesbania","ds80.shp"))

# Check the CRS:
sesbania %>% st_crs()

## Notice that this CRS is different from the California counties CRS, so we’ll want to update it to match. Use st_transform() to update the CRS:

sesbania <- st_transform(sesbania, 3857)

# Then check it: 
sesbania %>% st_crs()

## now they have the same CRS.
```
## Now plot them together
```{r}
ggplot() +
  geom_sf(data = ca_subset) +
  geom_sf(data = sesbania, size = 1, color = "red")
```
## more wrangling
-  say we want to find the count of red sesbania observed locations in this dataset by county.
- How can I go about joining these data so that I can find counts? Don’t worry…st_join() has it covered
```{r}
ca_sesbania <- ca_subset %>% 
  st_join(sesbania)

# And then we can find counts (note: these are not counts for individual plants, but by record in the dataset) by county:

sesbania_counts <- ca_sesbania %>% 
  count(county_name)

# Then we can plot a chloropleth using the number of records for red sesbania as the fill color (instead of what we used previously, land area):
ggplot(data = sesbania_counts) +
  geom_sf(aes(fill = n), color = "white", size = 0.1) +
  scale_fill_gradientn(colors = c("lightgray","orange","red")) +
  theme_minimal() +
  labs(fill = "Number of S. punicea records")
```
##  Only plot the county with the greatest number of red sesbania records (Solano), and make a map of those locations:
```{r}
# Subset of sesbania point locations only in Solano County
solano_sesbania <- sesbania %>% 
  filter(COUNTY == "Solano")

# Only keep Solano polygon from California County data
solano <- ca_subset %>% 
  filter(county_name == "Solano")

ggplot() +
  geom_sf(data = solano) +
  geom_sf(data = solano_sesbania)
```
## Make an interactive map with {tmap}
```{r}
# Set the viewing mode to "interactive":
tmap_mode(mode = "view")

# Then make a map (with the polygon fill color updated by variable 'land_area', updating the color palette to "BuGn"), then add another shape layer for the sesbania records (added as dots):
tm_shape(ca_subset) +
  tm_fill("land_area", palette = "BuGn") +
  tm_shape(sesbania) +
  tm_dots()
```

For extra resources see:
- https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html
- https://geocompr.robinlovelace.net/adv-map.html#interactive-maps
